### 2014-08-18 Pero_Transcriptome.tex
added figure 2
update

Deserts are widely considered one of Earth's harshest environments. Animals living in desert environments are forced to endure intense heat and drought, and in turn, species having evolved in these environments are likely ton have evolved specialised mechanisms that may enhance fitness. While living in deserts likely involves a large number of adaptive phenotypes, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}. Indeed, the maintenance of water balance in animals is one of the most important physiologic processes for all animals, desert inhabitants or not. Most animals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.  When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur, which suggests that there is strong selection for mechanisms supporting osmoregulation. Understanding these mechanisms will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, having implications for studies of human health, conservation, and climate change. \\

@@

The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers recommendations. The error corrected adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt. The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, I downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). I used a relaxed blastN procedure (evalue set to 10$^-10$) to identify contigs in the \textit{P. eremicus} dataset that are likely biological in origin. This procedure, when a reference dataset is available, is superior to a strategy employing expression-based filtering of the raw assembly. I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity. The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}). \\

@@

In addition to the reference individual sequenced at four different tissue types, we sequenced 15 other conspecific individuals from the same population, located in Palm Desert, California. Sequence data were mapped to the reference transcriptome using bwa-mem. The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd}, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \cite{Korneliussen:2013uz}. \\

@@

As detailed above, the RNAseq data from 15 individuals were mapped to the reference transcriptome with the resulting BAM files being used as input to the sortware package ANGSD. The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals. The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 1), with a much smaller fraction being subject to neutral or

>>>>

Deserts are widely considered one of Earth's harshest environments. Animals living in desert environments are forced to endure intense heat and drought, and in turn, species having evolved in these environments are likely to have evolved specialised mechanisms that may enhance fitness. While living in deserts likely involves a large number of adaptive phenotypes, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}. Indeed, the maintenance of water balance in animals is one of the most important physiologic processes for all animals, desert inhabitants or not. Most animals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.  When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur, which suggests that there is strong selection for mechanisms supporting osmoregulation. Understanding these mechanisms will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, having implications for studies of human health, conservation, and climate change. \\

@@

The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers defaultrecommendations. The error corrected adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt. The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, I downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). I used a blastN procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P. eremicus} dataset that are likely biological in origin. This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1) of the raw assembly. I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity. The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}). \\

@@

In addition to the reference individual sequenced at four different tissue types, we sequenced 15 other conspecific individuals from the same population, located in Palm Desert, California. Sequence data were mapped to the reference transcriptome using bwa-mem. The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd} version 0.610, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \cite{Korneliussen:2013uz}. \\

@@

As detailed above, the RNAseq data from 15 individuals were mapped to the reference transcriptome with the resulting BAM files being used as input to the software package ANGSD. The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals. The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 0), with a much smaller fraction being subject to neutral or positive selection.

\centerline{\includegraphics[width=18.0\baselineskip]{/Users/macmanes/Documents/pero_transcriptome/tajD.jpeg}}
\begin{quote}
\small{Figure 2. Tajima's D}
\end{quote}
